---
title: "How to make a data quality report"
author: "Edward Palmer"
date: "20th September 2019. Updated: `r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{dq_report}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The primary role of inspectEHR is to process the data inside CC-HIC in such a way as to make it ready to produce a data quality report.

First, install inspectEHR from github, or download the latest release directly.

```{r, eval = FALSE}
# If you need inspectEHR
library(devtools)
install_github("CC-HIC/inspectEHR")
```

You will most likely be working with sqlite as the postgres instance is confined to the UCL IDHS.

```{r, message = FALSE}
library(inspectEHR); library(dbplyr)

# Synthetic database ships with inspectEHR
#db_path <- system.file("testdata/synthetic_db.sqlite3", package = "inspectEHR")
#output_path <- "./2019-09-reports"
```

Supply the `report` function with an output folder. Please note, you must have write access to the database for this function to work.

```{r}
#report(sqlite_file = db_path, output_folder = output_path)
```

This will perform the following procedures:

1. Episodes will be characterised and verified/validated. This information will be stored inside the database in a new table called `episode_verification`.

2. Events will be verified/validated. This information will be stored inside the database in a new table called `event_verification`.

3. Summary information that is necessary for making a data quality report is exported as a `.RData` file to the `output_folder`.

4. Plots are pre-drawn and exported to the `output_folder`.

5. A data quality report is generated and exported to the `output_folder`.

6. A data quality score is generated and added to the report for each site.

Steps 3 and 4 are necessary as it allows for non identifying information to be exported from a secure location, without any risk of confidential information leak.
